---
title: "Analysis for Project"
output: 
  html_document: 
    fig_height: 6
    fig_width: 9
    toc: yes
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.align = "center")
source("useful_functions.R")
load("output/analytical_data.RData")
library(ggplot2)
library(scales)
library(texreg)
```

# Univariate distributions
<a href="#top">Back to top</a>

I created barplots to display the distributions of periods and feature types, as well as histograms of density. 

```{r period, echo=FALSE}
ggplot(huizui, aes(x=Period, y=..prop.., group=1))+
  geom_bar()+
  scale_y_continuous(labels = scales::percent)+
  labs(x="Periods", y=NULL,
       title = "Distribution of Periods")+
  theme_bw()
```

This barplot shows that there are three periods where the seeds are from--Yangshao, Longshan, and Erlitou. Most seed samples are from the Erlitou period (more than 80%). Yangshao and Longshan seeds constitute less than 10% of total seeds respectively, Yangshao has the smallest proportion among the three periods. 

```{r feature_type, echo=FALSE}
ggplot(huizui, aes(x=FeatureType, y=..prop.., group=1))+
  geom_bar()+
  scale_y_continuous(labels = scales::percent)+
  labs(x="Feature Types", y=NULL,
       title = "Distribution of Feature Types")+
  theme_bw()
```

This barplot shows that among the three feature types, ash pits are the most common one, representing more than half of all seed samples. Approximately 40% of seed samples are found in house context that is the second largest feature type. Slightly less than 15% of all seed samples are from ditches.

```{r density, echo=FALSE}
temp <- rbind(data.frame(density=huizui$TotalSeedDensity, seed_type="Total"),
              data.frame(density=huizui$MilletDensity, seed_type="Millet"),
              data.frame(density=huizui$LCDensity, seed_type="Luxury Cereals"),
              data.frame(density=huizui$BeanDensity, seed_type="Bean"))

ggplot(temp, aes(x=density))+
  geom_histogram(fill="gray", col = "black")+
  facet_wrap(~seed_type, scales = "free_x")+
  labs(x="density (#/L)",
       title="Histograms of Density")+
  theme_bw()
```

These four histograms show the densities of total seeds, millets, luxury cereals, and beans. Collectively, all four histograms have a minimum density of 0 #/L, which means there are soil samples with no seed/millet/luxury cereal/bean found. The histograms also have considerably different x scales. The total seed density histogram has a bin width of ~2 #/L. Most samples have total seed densities within the 0-20 #/L range, and the largest outlier reaches 60 #/L. The millet density histogram has a bin width of 1.25 #/L. Most samples with millets fall into the 0-10 #/L range, and the largest outlier has a millet density of more than 35 #/L. The luxury cereal density histogram has the lowest densities among the four. The bin width is ~0.07 #/L, and most samples have 0 or less than 0.07 #/L of LC densities. However, there's one sample with a LC density of more than 2.3 #/L, which is distinctively larger than other samples. The bean density histogram has the second lowest bin width of ~0.17 #/L. Most samples have bean densities within the 0-1 #/L range, and the largest outlier almost reaches 5 #/L. 

# Bivariate plots
<a href="#top">Back to top</a>

```{r Density_period, echo=FALSE}
temp1 <- rbind(data.frame(period=huizui$Period, density=huizui$TotalSeedDensity, seed_type="Total"),
              data.frame(period=huizui$Period, density=huizui$MilletDensity, seed_type="Millet"),
              data.frame(period=huizui$Period, density=huizui$LCDensity, seed_type="Luxury Cereals"),
              data.frame(period=huizui$Period, density=huizui$BeanDensity, seed_type="Bean"))

ggplot(temp1, aes(x=period, y=density, fill=period))+
  geom_boxplot()+
  facet_wrap(~seed_type, scales = "free_y")+
  labs(x="period", y="density (#/L)",
       title="Boxplots of Density by Period")+
  guides(fill=FALSE)+
  theme_bw()
```

These boxplots show the densities of total seed, millet, luxury cereals, and bean by period. The y (density) scales of the four plots are very different, which means millets are the most common, and there are more beans found among the samples than luxury cereals. The total seed density boxplot has the highest densities among the three periods than other boxplots. The Yangshao period samples have the lowest 1st quartile and median, while the Erlitou period has the highest 3rd quartile and most outliers. This is understandable because the Erlitou samples constitute the largest proportion among the three periods. This is also the case for the luxury cereals density boxplot and the bean density boxplot. However, the Yangshao period seems to have the highest millet densities among the three.

```{r Density_featureType, echo=FALSE}
temp2 <- rbind(data.frame(feature=huizui$FeatureType, density=huizui$TotalSeedDensity, seed_type="Total"),
              data.frame(feature=huizui$FeatureType, density=huizui$MilletDensity, seed_type="Millet"),
              data.frame(feature=huizui$FeatureType, density=huizui$LCDensity, seed_type="Luxury Cereals"),
              data.frame(feature=huizui$FeatureType, density=huizui$BeanDensity, seed_type="Bean"))

ggplot(temp2, aes(x=feature, y=density, fill=feature))+
  geom_boxplot()+
  facet_wrap(~seed_type, scales = "free_y")+
  labs(x="feature type", y="density (#/L)",
       title="Boxplots of Density by Feature Type")+
  guides(fill=FALSE)+
  theme_bw()
```

These boxplots show the densities of total seed, millet, luxury cereals, and bean by feature type. The y (density) scales of the four plots are very different, which means millets are the most common, and there are more beans found among the samples than luxury cereals. The total seed density boxplot has the highest densities among the three periods than other boxplots. House seems to be the feature from which most seeds especially millets are found. However, luxury cereals and beans are more commonly found in ash pits than in the other two feature types. 

# Poisson models
<a href="#top">Back to top</a>

```{r total_density_models, echo=FALSE, results="asis"}
total1 <- glm(TotalSeedNo~Period, family=quasipoisson, offset=log(Vol), data=huizui)

total2 <- glm(TotalSeedNo~FeatureType, family=quasipoisson, offset=log(Vol), data=huizui)

total3 <- glm(TotalSeedNo~Period+FeatureType, family=quasipoisson, offset=log(Vol), data=huizui)

total4 <- glm(TotalSeedNo~Period*FeatureType, family=quasipoisson, offset=log(Vol), data=huizui)

htmlreg(list(total1, total2, total3, total4),
          custom.coef.names = c("Intercept",
                             "Longshan Period",
                             "Yangshao Period",
                             "Ditch",
                             "House",
                             "Yangshao:Ditch",
                             "Longshan:House",
                             "Yangshao:House"),
          caption="Quasi-poisson models of total seed densities by period and feature types",
          caption.above = TRUE)
       
```

For total seed densities as the dependent variable, I created four quasi-poisson models with different independent variables. 

Model 1 has Period as the only independent variable. It shows that the Erlitou period soil samples have a total seed density of 12.93 #/L on average; the Longshan period samples have a total seed density of 29% less than Erlitou samples, on average; and the Yangshao period samples have a total seed density of 30% less than Erlitou samples, on average. Only the intercept is statistically significant. 

Model 2 has feature types as the only independent variable. It shows that soil samples from ash pits have a total seed density of 8.5 #/L on average; samples from ditches have a total seed density of 5% less than ash pit samples, on average; and samples from houses have a total seed density of 32% more than ash pit samples, on average. Only the intercept is statistically significant.

Model 3 has both perid and feature types as independent variables. It shows that Erlitou soil samples from ash pits have a total seed density of 12.18 #/L on average. Within the same feature type, the Longshan period samples have a total seed density of 29% less than Erlitou samples, on average; and the Yangshao period samples have a total seed density of 34% less than Erlitou samples, on average. Within the same period, samples from ditches have a total seed density of 9% less than ash pit samples, on average; and samples from houses have a total seed density of 35% more than ash pit samples, on average. Only the intercept is statistically significant.

Model 4 added the interaction between perid and feature types in addition to Model 3. It shows that Erlitou soil samples from ash pits have a total seed density of 13.6 #/L on average. Among samples from ash pits, the Longshan period samples have a total seed density of 22% less than Erlitou samples, on average; and the Yangshao period samples have a total seed density of 44% less than Erlitou samples, on average. Among the Erlitou samples, samples from ditches have a total seed density of 89% less than ash pit samples, on average; and samples from houses have a total seed density of 73% more than ash pit samples, on average. Among ditch samples, the total seed density from the Yangshao period is 28% more than that from the Erlitou period. Among house samples, the total seed density from the Longshan period is 74% less than that from the Erlitou period. Among house samples, the total seed density from the Yangshao period is 40% more than that from the Erlitou period. No coefficient in Model 4 is statistically significant. 

```{r millet_density_models, echo=FALSE, results="asis"}
millet1 <- glm(Millets~Period, family=quasipoisson, offset=log(Vol), data=huizui)

millet2 <- glm(Millets~FeatureType, family=quasipoisson, offset=log(Vol), data=huizui)

millet3 <- glm(Millets~Period+FeatureType, family=quasipoisson, offset=log(Vol), data=huizui)

millet4 <- glm(Millets~Period*FeatureType, family=quasipoisson, offset=log(Vol), data=huizui)

htmlreg(list(millet1, millet2, millet3, millet4),
          custom.coef.names = c("Intercept",
                             "Longshan Period",
                             "Yangshao Period",
                             "Ditch",
                             "House",
                             "Yangshao:Ditch",
                             "Longshan:House",
                             "Yangshao:House"),
          caption="Quasi-poisson models of millet densities by period and feature types",
          caption.above = TRUE)
       
```

For millet densities as the dependent variable, I created four quasi-poisson models with different independent variables. 

Model 1 has Period as the only independent variable. It shows that the Erlitou period soil samples have a millet density of 10.17 #/L on average; the Longshan period samples have a millet density of 55% less than Erlitou samples, on average; and the Yangshao period samples have a millet density of 69% less than Erlitou samples, on average. Only the intercept and the coefficient of Yangshao period are statistically significant. 

Model 2 has feature types as the only independent variable. It shows that soil samples from ash pits have a millet density of 3.42 #/L on average; samples from ditches have a millet density of 35% less than ash pit samples, on average; and samples from houses have a millet density of 49% more than ash pit samples, on average. Only the intercept is statistically significant.

Model 3 has both perid and feature types as independent variables. It shows that Erlitou soil samples from ash pits have a millet density of 9.77 #/L on average. Within the same feature type, the Longshan period samples have a millet density of 57% less than Erlitou samples, on average; and the Yangshao period samples have a millet density of 93% less than Erlitou samples, on average. Within the same period, samples from ditches have a millet density of 46% less than ash pit samples, on average; and samples from houses have a millet density of 61% more than ash pit samples, on average. Only the intercept and the coefficient of Yangshao period are statistically significant. 

Model 4 added the interaction between perid and feature types in addition to Model 3. It shows that Erlitou soil samples from ash pits have a millet density of 11.47 #/L on average. Among samples from ash pits, the Longshan period samples have a millet density of 56% less than Erlitou samples, on average; and the Yangshao period samples have a millet density of 81% less than Erlitou samples, on average. Among the Erlitou samples, samples from ditches have a millet density of 95% less than ash pit samples, on average; and samples from houses have a millet density of 50% more than ash pit samples, on average. Among ditch samples, the millet density from the Yangshao period is 14% more than that from the Erlitou period. Among house samples, the millet density from the Longshan period is 58% less than that from the Erlitou period. Among house samples, the millet density from the Yangshao period is 91% more than that from the Erlitou period. No coefficient in Model 4 is statistically significant.

```{r LC_density_models, echo=FALSE, results="asis"}
LC1 <- glm(LuxuryCereal~Period, family=quasipoisson, offset=log(Vol), data=huizui)

LC2 <- glm(LuxuryCereal~FeatureType, family=quasipoisson, offset=log(Vol), data=huizui)

LC3 <- glm(LuxuryCereal~Period+FeatureType, family=quasipoisson, offset=log(Vol), data=huizui)

LC4 <- glm(LuxuryCereal~Period*FeatureType, family=quasipoisson, offset=log(Vol), data=huizui)

htmlreg(list(LC1, LC2, LC3, LC4),
          custom.coef.names = c("Intercept",
                             "Longshan Period",
                             "Yangshao Period",
                             "Ditch",
                             "House",
                             "Yangshao:Ditch",
                             "Longshan:House",
                             "Yangshao:House"),
          caption="Quasi-poisson models of luxury cereal densities by period and feature types",
          caption.above = TRUE)
       
```

For luxury cereal (LC) densities as the dependent variable, I created four quasi-poisson models with different independent variables. 

Model 1 has Period as the only independent variable. It shows that the Erlitou period soil samples have a LC density of 0.04 #/L on average; the Longshan period samples have a LC density of 400% more than Erlitou samples, on average; and the Yangshao period samples have a LC density of 116% more than Erlitou samples, on average. Only the intercept is statistically significant. 

Model 2 has feature types as the only independent variable. It shows that soil samples from ash pits have a LC density of 0.13 #/L on average; samples from ditches have a LC density of 46% less than ash pit samples, on average; and samples from houses have a LC density of 43% less than ash pit samples, on average. Only the intercept is statistically significant.

Model 3 has both perid and feature types as independent variables. It shows that Erlitou soil samples from ash pits have a LC density of 0.06 #/L on average. Within the same feature type, the Longshan period samples have a LC density of 335% more than Erlitou samples, on average; and the Yangshao period samples have a LC density of 116% more than Erlitou samples, on average. Within the same period, samples from ditches have a LC density of 34% less than ash pit samples, on average; and samples from houses have a LC density of 39% less than ash pit samples, on average. Only the intercept is statistically significant.

Model 4 added the interaction between perid and feature types in addition to Model 3. It shows that Erlitou soil samples from ash pits have a LC density of 0.1 #/L on average. Among samples from ash pits, the Longshan period samples have a LC density of 189% more than Erlitou samples, on average; and the Yangshao period samples have a LC density of 17% more than Erlitou samples, on average. Among the Erlitou samples, samples from ditches have an almost identical LC density as ash pit samples, on average; and samples from houses also have the same LC density as ash pit samples, on average. Among ditch samples, the LC density from the Yangshao period is 20% less than that from the Erlitou period. Among house samples, the LC density from the Longshan period is the same as that from the Erlitou period. Among house samples, the LC density from the Yangshao period is 26% less than that from the Erlitou period. Only the intercept is statistically significant.

```{r bean_density_models, echo=FALSE, results="asis"}
bean1 <- glm(Bean~Period, family=quasipoisson, offset=log(Vol), data=huizui)

bean2 <- glm(Bean~FeatureType, family=quasipoisson, offset=log(Vol), data=huizui)

bean3 <- glm(Bean~Period+FeatureType, family=quasipoisson, offset=log(Vol), data=huizui)

bean4 <- glm(Bean~Period*FeatureType, family=quasipoisson, offset=log(Vol), data=huizui)

htmlreg(list(bean1, bean2, bean3, bean4),
          custom.coef.names = c("Intercept",
                             "Longshan Period",
                             "Yangshao Period",
                             "Ditch",
                             "House",
                             "Yangshao:Ditch",
                             "Longshan:House",
                             "Yangshao:House"),
          caption="Quasi-poisson models of bean densities by period and feature types",
          caption.above = TRUE)
       
```

For bean densities as the dependent variable, I created four quasi-poisson models with different independent variables. 

Model 1 has Period as the only independent variable. It shows that the Erlitou period soil samples have a bean density of 0.06 #/L on average; the Longshan period samples have a bean density of 274% more than Erlitou samples, on average; and the Yangshao period samples have a bean density of 569% more than Erlitou samples, on average. No coefficient is statistically significant. 

Model 2 has feature types as the only independent variable. It shows that soil samples from ash pits have a bean density of 0.51 #/L on average; samples from ditches have a bean density of 23% less than ash pit samples, on average; and samples from houses have a bean density of 71% less than ash pit samples, on average. Only the intercept and the coefficient of House are statistically significant.

Model 3 has both perid and feature types as independent variables. It shows that Erlitou soil samples from ash pits have a bean density of 0.08 #/L on average. Within the same feature type, the Longshan period samples have a bean density of 225% more than Erlitou samples, on average; and the Yangshao period samples have a bean density of 639% more than Erlitou samples, on average. Within the same period, samples from ditches have a bean density of 19% less than ash pit samples, on average; and samples from houses have a bean density of 74% less than ash pit samples, on average. Only the coefficient of House is statistically significant. 

Model 4 added the interaction between perid and feature types in addition to Model 3. It shows that Erlitou soil samples from ash pits have a bean density of 0.06 #/L on average. Among samples from ash pits, the Longshan period samples have a bean density of 267% more than Erlitou samples, on average; and the Yangshao period samples have a bean density of 839% more than Erlitou samples, on average. Among the Erlitou samples, samples from ditches have a bean density of 9% less than ash pit samples, on average; and samples from houses have a bean density of 9% less than ash pit samples, on average. Among ditch samples, the bean density from the Yangshao period is 20% less than that from the Erlitou period. Among house samples, the bean density from the Longshan period is 16% less than that from the Erlitou period. Among house samples, the bean density from the Yangshao period is 75% less than that from the Erlitou period. No coefficient in Model 4 is statistically significant.