---
title: "Analysis for Project"
output: 
  html_document: 
    fig_height: 6
    fig_width: 9
    toc: yes
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.align = "center")
source("useful_functions.R")
load("output/analytical_data.RData")
library(ggplot2)
library(texreg)
```

# Add several new variables

```{r densities, include=FALSE}
huizui$MilletDensity <- huizui$Millets / huizui$Vol
huizui$LCDensity <- huizui$LuxuryCereal / huizui$Vol
huizui$BeanDensity <- huizui$Bean / huizui$Vol
```

# Univariate distributions

```{r period, fig.cap="Distribution of period"}
ggplot(huizui, aes(x=Period, y=..prop.., group=1))+
  geom_bar()+
  theme_bw()
```

```{r feature_type, fig.cap="Distribution of Feature Type"}
ggplot(huizui, aes(x=FeatureType, y=..prop.., group=1))+
  geom_bar()+
  theme_bw()
```

```{r total_seed_density, fig.cap="Boxplot of Total Seed Density"}
boxplot(huizui$TotalSeedDensity, col = "royalblue")
```

```{r millet_density, fig.cap="Boxplot of Millet Density"}
boxplot(huizui$MilletDensity, col = "royalblue")
```

```{r LC_density, fig.cap="Boxplot of Luxury Cereal Density"}
boxplot(huizui$LCDensity, col = "royalblue")
```

```{r bean_density, fig.cap="Boxplot of Bean Density"}
boxplot(huizui$BeanDensity, col = "royalblue")
```

# Bivariate plots

```{r totalDensity_period, echo=FALSE}
boxplot(huizui$TotalSeedDensity ~ huizui$Period, main = "Total Seed Density by Periods", xlab = "Periods", ylab = "Total Seed Density (#/L)", col = rainbow(4))
```

```{r totalDensity_featureType, echo=FALSE}
boxplot(huizui$TotalSeedDensity ~ huizui$FeatureType, main = "Total Seed Density by Feature Types", xlab = "Feature Types", ylab = "Total Seed Density (#/L)", col = rainbow(6))
```


# Introduction
<a href="#top">Back to top</a>

Use this R Markdown to perform the main analysis for the project. I use this basically as a lab notebook. It contains the main analysis and a variety of sensitivity analysis. The code in this documents serves as a baseline for the eventual tables and figures that will go into the paper. At the same time it will serve as a record of all supplementary analyses performed. 